Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL

### 2020-12-07 main.tex
Updates from Overleaf
- paper changes

	\begin{abstract}
		In this paper, we compare a model-based reinforcement learning to a model-free reinforcement learning method applied to the FERMI FEL system. Both approaches hold tremendous promise, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level in a reasonable training time on real accelerator physics problems. In terms of representational power and sample-efficiency, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior. The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning. In both cases, the techniques were adapted to show increased noise resistivity as omnipresent in accelerator control problems. Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.

@@

	 In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}. To train these methods, a large number of direct interactions with the system is required, because they belong to on-policy algorithms. In the latter one deals with approximate dynamic programming, which we shortly treat in \cref{s:Model free reinforcement learning}. One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained. A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010}.

@@

		Q^\policy(\bs,\ba) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],

@@

	A nice overview of model-based reinforcement learning (MBRL) methods is provided in \cite{Wang2019}. Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods. We focus on model-based data generation.\\

@@

Several methods of network training were implemented. Early stopping, with a learning rate of $10^{-3}$ to $10^{-4}$  was employed  with a waiting time of about 20 steps. The number of training steps is increasing with more data in close-by areas. It is followed by a shrinking of the uncertainty at regions where no data is available, leading to better training performance. A fixed maximal loss threshold was also beneficially tested. In the experiments, a combination of both was taken.\\

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference and code clean up.
	\item Conclusion and outlook\dots
	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
	\item Check the plots.
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}

>>>>

	\begin{abstract}
		In this paper, we compare a model-based reinforcement learning to a model-free reinforcement learning method applied to the FERMI FEL system. Both approaches hold tremendous promise, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level in a reasonable training time on real accelerator physics problems. In terms of representational power and sample-efficiency, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior. The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning. In both cases, the techniques were adapted to show increased noise robustness as omnipresent in accelerator control problems. Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.

@@

	 In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}. To train these methods, a large number of direct interactions with the system is required, because they belong to on-policy algorithms. In the latter one deals with approximate dynamic programming, which we shortly treat in \cref{s:Model free reinforcement learning}. One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained. A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010,Lillicrap2015,pmlr-v32-silver14}.

@@

		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],

@@

	A nice overview of model-based reinforcement learning (MBRL) methods is provided in \cite{Wang2019}. Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods (e.g. the cross entropy method \cite{Boer2005}). We focus on model-based data generation.\\

@@

Several methods of network training were implemented. \emph{Early stopping} \cite{Goodfellow2016}, with a learning rate of $10^{-3}$ to $10^{-4}$  was employed  with a waiting time of about 20 steps and a validation ratio of 20\%. The number of training steps is increasing with more data in close-by areas. It is followed by a shrinking of the uncertainty at regions where no data is available, leading to better training performance. A fixed maximal loss threshold was also beneficially tested. In the experiments, a combination of both was taken.\\

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference and code clean up.
	\item Conclusion and outlook\dots
	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
	\item Check the plots, which might be already okay!
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}
